Bird beak's elaboration and exploration

### 2021-05-26 Analysis/02-MCMCglmm_mini_chains.Rmd
Update 02-MCMCglmm_mini_chains.Rmd

```{r load_beer, eval = TRUE}
## Loading beer
library(beer)

## Loading the demo data
data(demo_data)
data(demo_trees)
```

@@

We first ran three independent MCMC chains with the model and data described above using three random trees and flat priors.
We ran these models (for `§§§N_generations` with sampling every `§§§N_nitt` and no burnin) until the three chains reached an effective sample size >> 200 (§§§and other diagnosis).
From these three chains, we first extracted the three burnin periods (defined as the generation when the chain reaches the median likelihood value times 1.25).
We then extracted the posterior estimates of the R-Structure, G-Structure and fixed effects (mean and covariance) (ignoring the runs from the previously estimated burnin periods) to serve as our priors for our mini-chains. 
For those values, we took the highest burnin as the overall burnin and the median of the three posteriors as the overall priors.

```{r, parametrise_the_model}
## Set up the parametrising chains models
param_MCMCglmm <- make.mini.chains(data         = demo_data,
                                   dimensions   = c(1:4),
                                   tree         = demo_trees,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "clade"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 10000,
                                       thin   = 100,
                                       burnin = 0))

@@

@@

We set up a mini-chain to be a MCMCglmm running with 1) the model described above, 2) a random tree and 3) the estimated priors (with a degree of parameter belief of 5% `nu = 0.05`) to run for `nitt` generations were `nitt` is equal to the previously estimated burnin phase + 1000 sampled iterations (`nitt = burnin + thin * 1000`).
Each mini-chain thus provided us with a 1000 exploitable posterior data points.
The two main advantages of this mini-chains is that 1) they are much faster to run since no diagnoses is necessary and the chains are only ran for a relatively short time (which allow several chains to crash/fail without loosing all the data); and 2) they take into account tree uncertainty whitout having to run the complete MCMCglmm on all trees (c.f. @mulTree).

```{r, parametrise_the_model}
## Set up the mini-chains models

@@

>>>>

```{r load_beer, eval = TRUE}
set.seed(42)
## Loading beer
library(beer)

## Loading the demo data
data(demo_data)
data(demo_trees)
data(demo_tree)

## Setting the consensus tree
consensus_tree <- demo_tree[[1]]
```

@@

We first ran three independent MCMC chains with the model and data described above using the consensus tree and flat priors.
We ran these models (for `§§§N_generations` with sampling every `§§§N_nitt` and no burnin) until the three chains reached an effective sample size >> 200 (§§§and other diagnosis).

```{r, parametrise_the_model}
## Set up the parametrising chains models
param_MCMCglmm <- make.mini.chains(data         = demo_data,
                                   dimensions   = c(1:4),
                                   tree         = consensus_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "clade"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 10000,
                                       thin   = 100,
                                       burnin = 0))

@@

From these three chains, we first extracted the three burnin periods (defined as the generation when the chain reaches the median likelihood value times 1.25).
We then extracted the posterior estimates of the R-Structure, G-Structure and fixed effects (mean and covariance) (ignoring the runs from the previously estimated burnin periods) to serve as our priors for our mini-chains. 
For those values, we took the highest burnin as the overall burnin and the median of the three posteriors as the overall priors.

```{r, get_model_parameters}
## Extract the parameters from the parameter chains
est_params <- extract.parameters(param_chains)
```

@@

We set up a mini-chain to be a MCMCglmm running with 1) the model described above, 2) a random tree and 3) the estimated priors (with a degree of parameter belief of 5% `nu = 0.05`) to run for `nitt` generations were `nitt` is equal to the previously estimated burnin phase + 100 sampled iterations (`nitt = burnin + thin * 100`).
Each mini-chain thus provided us with a 100 exploitable posterior data points.
The two main advantages of this mini-chains is that 1) they are much faster to run since no diagnoses is necessary and the chains are only ran for a relatively short time (which allow several chains to crash/fail without loosing all the data); and 2) they take into account tree uncertainty whitout having to run the complete MCMCglmm on all trees (c.f. @mulTree).

```{r, parametrise_the_model}
## The thining parameter
thin <- 100
## The number of samples per chains
samples <- 100

## Set up the mini-chains models
mini_chains <- make.mini.chains(data         = demo_data,
                                dimensions   = c(1:4),
                                tree         = demo_trees,
                                trait.family = "gaussian",
                                randoms      = c("global", "clade"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                    nitt   = est_params$burnin + thin * samples,
                                    thin   = thin,
                                    burnin = est_params$burnin))
```

We then run the MCMCglmm over $N_trees$ to get an estimate of the variance co-variance matrices including phylogenetic uncertainty.

```{r, running_replicates}
## The number of replicates
n_reps <- 10

## Run the required chains
results <- run.mini.chains(mini_chains, replicates = n_reps, path = "../Data/Processed/MCMC/")
```

#### Number of runs optimisation (TO BE DEVELOPPED?)

```{r, eval = FALSE}
## Function for running one mini chain on a random tree (and saving the results)
fun.mini.chains <- function() {
    run.mini.chains(mini_chains, replicates = 1, path = "../Data/Processed/MCMC/", file.name = "mini.chains")
}
```

@@

```{r, optimised_replicates, eval = FALSE}
## The summary function
